---
title: "Johns Hopkins Data Science Capstone Final Project"
author: "Jason Parise"
date: "`r Sys.Date()`"
output: html_document
---

### Johns Hopkins Data Science Capstone Final Project
### WORD PREDICTOR GENIE!

## The following code creates required ngram data files used to predict ngrams.
## Use these generated files in conjunction with server.R and ui.R files.
## The application will predict the next word of phrases 1-5 letters long.


# Required packages. NOTE: This must be used with R version 4.1.3.

```{r packages}
      library(tidytext)
      library(tidyverse)
      library(stringr)
      library(knitr)
      library(wordcloud)
      library(ngram)
      library(shiny)
```

# Download and unzip English language data.
```{r download}
download.file("https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip", "Coursera-SwiftKey.zip")
unzip("Coursera-SwiftKey.zip")
```

# Load and read the data.
```{r data}
blogs <- "./final/en_US/en_US.blogs.txt"
news <- "./final/en_US/en_US.news.txt"
twitters <- "./final/en_US/en_US.twitter.txt"
blogs <- readLines(blogs, skipNul = TRUE)
con <- file(news, open="rb")
news <- readLines(con,  skipNul = TRUE)
twitters <- readLines(twitters, skipNul = TRUE)
```

# Reformat data into more useful dataframes.
```{r reform}
blogs <- tibble(text = blogs)
news <- tibble(text = news)
twitters <- tibble(text = twitters)
```

# Recreating a sampling of all the data from 2.5% of the bulk.
```{r sample}
set.seed(25)
sample_percentage <- 0.025
blog1 <- blogs %>%
      sample_n(., nrow(blogs)*sample_percentage)
news1 <- news %>%
      sample_n(., nrow(news)*sample_percentage)
twitter1 <- twitters %>%
      sample_n(., nrow(twitters)*sample_percentage)
```

# Making data clean and tidy.
```{r clean}
TidyData <- bind_rows(
      mutate(blog1, source = "blogs"),
      mutate(news1,  source = "news"),
      mutate(twitter1, source = "twitters"))
TidyData$source <- as.factor(TidyData$source)
data("stop_words")
space <- "[^[:alpha:][:space:]]*"
urls <- "http[^[:space:]]*"
character <- "\\b(?=\\w*(\\w)\\1)\\w+\\b"
CleanTidyData <-  TidyData %>%
      mutate(text = str_replace_all(text, space, "")) %>%
      mutate(text = str_replace_all(text, urls, "")) %>%
      mutate(text = str_replace_all(text, character, "")) %>%
      mutate(text = iconv(text, "ASCII//TRANSLIT"))
```

# Ngram generation.
```{r ngram}
Unigram <- CleanTidyData %>%
      unnest_tokens(word, text) %>%
      anti_join(stop_words)

Bigram <- CleanTidyData %>%
      unnest_tokens(bigram, text, token = "ngrams", n = 2)

Trigram <- CleanTidyData %>%
      unnest_tokens(trigram, text, token = "ngrams", n = 3)

Quadgram <- CleanTidyData %>%
      unnest_tokens(quadgram, text, token = "ngrams", n = 4)

Quintgram <- CleanTidyData %>%
      unnest_tokens(quintgram, text, token = "ngrams", n = 5)

Sextgram <- CleanTidyData %>%
      unnest_tokens(sextgram, text, token = "ngrams", n = 6)

Bigram1 <- Bigram %>%
      count(bigram) %>%
      filter(n > 10) %>%
      arrange(desc(n))

Trigram1 <- Trigram %>%
      count(trigram) %>%
      filter(n > 10) %>%
      arrange(desc(n))
      
Quadgram1 <- Quadgram %>%
      count(quadgram) %>%
      filter(n > 10) %>%
      arrange(desc(n))
      
Quintgram1 <- Quintgram %>%
      count(quintgram) %>%
      filter(n > 10) %>%
      arrange(desc(n))

Sextgram1 <- Sextgram %>%
      count(sextgram) %>%
      filter(n > 10) %>%
      arrange(desc(n))
      
Bigram2 <- Bigram1 %>%
      separate(bigram, c("word1", "word2"), sep = " ")

Trigram2 <- Trigram1 %>%
      separate(trigram, c("word1", "word2", "word3"), sep = " ")

Quadgram2 <- Quadgram1 %>%
      separate(quadgram, c("word1", "word2", "word3", "word4"), sep = " ")

Quintgram2 <- Quintgram1 %>%
      separate(quintgram, c("word1", "word2", "word3", "word4", "word5"), sep = " ")

Sextgram2 <- Sextgram1 %>%
      separate(sextgram, c("word1", "word2", "word3", "word4", "word5", "word6"), sep = " ")
```

# Save all ngrams for use in the app.
```{r save}
dir.create("ngrams")

saveRDS(Bigram2, "./ngrams/bigrams.rds")
saveRDS(Trigram2, "./ngrams/trigrams.rds")
saveRDS(Quadgram2,"./ngrams/quadgrams.rds")
saveRDS(Quintgram2,"./ngrams/quintgrams.rds")
saveRDS(Sextgram2,"./ngrams/sextgrams.rds")
```
